---
title: "Week12_Assignment"
author: "Catherine Broski"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment

### Purpose

The goal of this assignment is to practice writing and using for loops for iteration.

### Task

Write R code to successfully answer each question below.

### Criteria for Success

-   Code is within the provided code chunks or new code chunks are created where necessary
-   Code chunks run without errors
-   Code chunks have brief comments indicating which code is answering which part of the question
-   Code will be assessed as follows:
    -   Produces the correct answer using the requested approach: 100%
    -   Generally uses the right approach, but a minor mistake results in an incorrect answer: 90%
    -   Attempts to solve the problem and makes some progress using the core concept, but returns the wrong answer and does not demonstrate comfort with the core concept: 50%
    -   Answer demonstrates a lack of understanding of the core concept: 0%
-   Any questions requiring written answers are answered with sufficient detail

### Due Date

April 15 at midnight MST

# Assignment Exercises

### Set Up

Be sure to load the `tidyverse` to use later in the assignment.

```{r}
library(tidyverse)
```

### 1. For Loop Basics (30 pts)

Complete the following tasks.

I recommend copying the commented code and making changes to the copy. That way, if you have made edits that you haven't kept track of, you have the original code to start over as needed.

a.  The code below prints the numbers 1 through 5 one line at a time. Modify it to print each of these numbers multiplied by 3.

```{r}
numbers <- c(1, 2, 3, 4, 5)
# for (number in numbers){
#   print(number)
# }

 for (number in numbers){
   number <- number * 3
   print(number)
 }
```

b.  Write a for loop that loops over the following vector and prints out the mass in kilograms (`mass_kg = 2.2 * mass_lb`)

```{r}
mass_lbs <- c(2.2, 3.5, 9.6, 1.2)

for (mass_lb in mass_lbs) {
  mass_kg <- 2.2 * mass_lb
  print(mass_kg)
}
```

c.  Complete the code below so that it prints out the name of each bird one line at a time.

```{r}
birds = c('robin', 'woodpecker', 'blue jay', 'sparrow')
for (i in 1:length(birds)){
   print(birds[i])
 }
```

d.  Complete the code below so that it stores one area for each radius.

```{r}
radius <- c(1.3, 2.1, 3.5)
 areas <- vector(mode = "numeric", length = length(radius))
 for (i in 1:length(radius)){
   areas[i] <- pi * radius[i] ^ 2
 }
areas
```

e.  Complete the code below to calculate an area for each pair of `lengths` and `widths`, store the areas in a vector, and after they are all calculated print them out:

```{r}
lengths = c(1.1, 2.2, 1.6)
widths = c(3.5, 2.4, 2.8)
 areas <- vector(mode = "numeric", length = length(lengths))
 for (i in 1:length(lengths)) {
   area <- lengths[i] * widths[i]
   areas[i] <- area
 }
areas
```

### 2. Size Estimates by Name (30 pts)

This is a followup to "Size Estimates by Name" from last week.

Download the `dinosaur_lengths.csv` file from D2L and place it in the correct directory. Read the file into R.

Write a function `mass_from_length()` that uses the equation `mass <- a * length^b` to estimate the size of a dinosaur from its length. This function should take two arguments, `length` and `species`. For each of the following inputs for `species`, so use the associated `a` and `b` values to estimate the species mass using these equations:

-   *Stegosauria*: `mass = 10.95 * length ^ 2.64` ([Seebacher 2001](http://www.jstor.org/stable/4524171)).
-   *Theropoda*: `mass = 0.73 * length ^ 3.63` ([Seebacher 2001](http://www.jstor.org/stable/4524171)).
-   *Sauropoda*: `mass = 214.44 * length ^ 1.46` ([Seebacher 2001](http://www.jstor.org/stable/4524171)).
-   For any other value of `species`: `mass = 25.37 * length ^ 2.49`

a.  Use this function and a for loop to calculate the estimated mass for each dinosaur, store the masses in a vector, and after all of the calculations are complete show the first few items in the vector using `head()`.
b.  Add the results in the vector back to the original data frame. Show the first few rows of the data frame using `head()`.
c.  Calculate the mean mass for each `species` using `dplyr` (no `for` loops).

```{r}
dino_lengths <- read_csv("../data_raw/dinosaur_lengths.csv")

mass_from_length <- function(length, species) {
  if (species == "Stegosauria") {
    mass = 10.95 * length ^ 2.64
  } else if (species == "Theropoda") {
    mass = 0.73 * length ^ 3.63 
  } else if (species == "Sauropoda") {
    mass = 214.44 * length ^ 1.46
  } else {
    mass = 25.37 * length ^ 2.49
  }
  return(mass)
}

# 2a

masses <- vector(mode = "numeric", length = nrow(dino_lengths)) 
  
for (i in 1:nrow(dino_lengths)) {
  mass <- mass_from_length(dino_lengths[i, "lengths"],
                           dino_lengths[i, "species"])
  masses[i] <- mass
}
 head(masses)

 # running, but returning a vertical list instead of horizontal
 # change to method with dollar signs, but welcome to leave it
 
 
 # 2b
 
 dino_lengths$mass <- masses 
 
 # for (i in 1:length(masses)) {
 # dino_lengths$mass[i] <- masses[i]
 #}
 head(dino_lengths)
 
 # mass column is added, but not showing observations when using head()
 
 # 2c
 
 #summarize
```

```{r}

```

### 3. Multi-file Analysis (40 pts)

You have satellite collars on a number of different individuals and want to be able to quickly look at all of their recent movements at once. The data is posted daily to a URL that contains one csv file for each individual: [zip file](http://www.datacarpentry.org/semester-biology/data/individual_collar_data.zip)

Start your solution by:

-   If `individual_collar_data.zip` is not already in your working directory, download it from the link above and place the zip file in the correct directory.
-   Unzip it using `unzip()`
-   Obtain a list of all of the files with file names matching the pattern `"collar-data-.*.txt"` (using `list.files()`)

a.  Use a loop to load each of these files into R and make a line plot (using `geom_path()`) for each file with `long` on the `x` axis and `lat` on the `y` axis.

Graphs, like other types of output, won't display inside a loop unless you explicitly display them, so you need put your `ggplot()` command inside a `print()` statement.

Include the name of the file in the graph as the graph title using the `ggtitle()`.

b.  Add code to the loop to calculate the minimum and maximum latitude in the file, and store these values, along with the name of the file, in a data frame.

Show the data frame as output.

```{r}
unzip("..data_raw/individual_collar_data.zip")

list.files(path = "../data_raw")

collar_files <- list.files(path = "../data_raw/individual_collar_data/",
                         pattern = "collar-data-.*.txt")
collar_files

# 3a
for (i in 1:length(collar_files)) {
  filename <- collar_files[i]
  data <- read.csv(paste0("../data_raw/individual_collar_data/", filename)) 
  plot <- ggplot(data = data, mapping = aes(x = long, y = lat)) +
    geom_path()
  print(plot)
}



```

# Reflection

This reflection is worth 25 points, separate from your Week 12 assignment. Reflections are graded for completion only.

Write about 5 sentences addressing *at least* one of the following questions.

-   What has worked well for you in this course for you so far?
-   What has been particularly challenging for you so far?
-   Is there anything that I can do to help your learning in the course?
-   Other reflections about the content of the course that you would like to share.

*Answer*:

Everything in the course is going well for me. I feel like I am catching on to concepts quickly and the assignments are not taking me too long to complete. I appreciate how receptive Ellen has been to my questions over email. I am nervous about the final project because of my workload outside of this class, but it will be helpful to have time to complete the project after the last day of classes. I'm looking forward to using data that is more similar to my field of study in the final project.
